Modelling the Lymphatic Metastatic Progression Pathways of OPSCC from Multi-Institutional Datasets

### 2022-12-05 src/tex/ms.tex
[showyourwork] overleaf sync

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}
\usepackage{multirow}

@@

The transition from one time-step to another is governed by the transition probability $P\left( \mathbf{X}[t+1]=\boldsymbol{\xi}_i \mid \mathbf{X}[t]=\boldsymbol{\xi}_j \right)$, which can conveniently be collected into a transition matrix when we enumerate all $2^V$ distinct possible states $\boldsymbol{\xi}_i$ with $i \in \left\{ 1,2, \ldots, 2^V \right\}$ of lymphatic involvement:
%
\begin{equation}
    \mathbf{A} = \left( A_{ij} \right) = \left( P\left( \mathbf{X}[t+1]=\boldsymbol{\xi}_i \mid \mathbf{X}[t]=\boldsymbol{\xi}_j \right) \right)
\end{equation}
%
The term inside the brackets describes the probability to transition from the hidden state of lymphatic involvement $\boldsymbol{\xi}_j$ to the state $\boldsymbol{\xi}_i$ within between the time $t$ and $t+1$. We can use a \gls{dag} as depicted in

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/graph_with_obs.png}
    \caption{\Gls{dag} representing a possible abstraction of the lymphatic network comprising the tumor (red shaded circle) and \glspl{lnl} I through IV as hidden binary \glspl{rv} (blue outlined circles). Attached to each of these is the corresponding observed \gls{rv} (orange shaded squares). Lymphatic flow is depicted in the form of parameterized arrows (red and blue) that represent the probability of spread along the respective arc per time-step. Sensitivity and specificity connect the hidden \glspl{rv} to the diagnosis (orange arrows).}
    \label{fig:graph_with_obs}
\end{figure}

@@

\begin{center}
\begin{tabular}{|cc|cc|}
    \hline
    & & \multicolumn{2}{c|}{$X$} \\
    & & 0 & 1 \\
    \hline
    \multirow{2}{*}{$Z$} & 0 & $s_p$ & $1 - s_N$ \\
    & 1 & $1 - s_P$ & $s_N$ \\
    \hline
\end{tabular}
\end{center}

>>>>

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}
\usepackage{multirow}
\usepackage{hyperref}
\usepackage{cleveref}

@@

The transition from one time-step to another is governed by the transition probability $P\left( \mathbf{X}[t+1]=\boldsymbol{\xi}_i \mid \mathbf{X}[t]=\boldsymbol{\xi}_j \right)$, which can conveniently be collected into a transition matrix when we enumerate all $2^V$ distinct possible states $\boldsymbol{\xi}_i$ with $i \in \left\{ 1,2, \ldots, 2^V \right\}$ of lymphatic involvement:
%
\begin{equation}
    \mathbf{A} = \left( A_{ij} \right) = \left( P\left( \mathbf{X}[t+1]=\boldsymbol{\xi}_i \mid \mathbf{X}[t]=\boldsymbol{\xi}_j \right) \right)
\end{equation}
%
The term $P\left( \boldsymbol{\xi}_i \mid \boldsymbol{\xi}_j \right)$ describes the probability to transition from the hidden state of lymphatic involvement $\boldsymbol{\xi}_j$ to the state $\boldsymbol{\xi}_i$ within between the time $t$ and $t+1$. Using a \gls{dag} as depicted in \cref{fig:graph_with_obs}, we can parameterize this transition probability in the following way:
%
\begin{equation}
    P\left( \boldsymbol{\xi}_i \mid \boldsymbol{\xi}_j \right) = \prod_{v \leq V} Q\left( \xi_{iv} ; \xi_{jv} \right) P \left( \xi_{iv} \mid \left\{ \xi_{jr} \right\}_{r \in \operatorname{pa}(v)} \right)^{1 - \xi_{jv}}
\end{equation}
%
In this equation, we have denoted \glspl{lnl} that are parents of \gls{lnl} $v$ with the symbol $r\in\operatorname{pa}(v)$. Also, $\xi_{iv}$ denotes the value that \gls{lnl} $v$ takes on when the patient is in state $\boldsymbol{\xi}_i$.
\begin{figure}
    \centering
    \includegraphics[width=0.6\linewidth]{figures/graph_with_obs.png}
    \caption{\Gls{dag} representing a possible abstraction of the lymphatic network comprising the tumor (red shaded circle) and \glspl{lnl} I through IV as hidden binary \glspl{rv} (blue outlined circles). Attached to each of these is the corresponding observed \gls{rv} (orange shaded squares). Lymphatic flow is depicted in the form of parameterized arrows (red and blue) that represent the probability of spread along the respective arc per time-step. Sensitivity and specificity (orange arrows) connect the hidden \glspl{rv} to the diagnosis.}
    \label{fig:graph_with_obs}
\end{figure}

@@

\begin{table}[h!]
    \centering
    \begin{tabular}{|cc|cc|}
        \hline
        & & \multicolumn{2}{c|}{$X$} \\
        & & 0 & 1 \\
        \hline
        \multirow{2}{*}{$Z$} & 0 & $s_p$ & $1 - s_N$ \\
        & 1 & $1 - s_P$ & $s_N$ \\
        \hline
    \end{tabular}
\end{table}
